Modeling thermal dust emission with two components: application to the {\it PLANCK} HFI maps

### 2014-07-22 ms.tex
audited for FDS99-like versus FDS99

\title{Two-component thermal dust emission model: application to the 
{\it PLANCK} HFI maps}
\author{Aaron M. Meisner\altaffilmark{1,2}}
\author{Douglas P. Finkbeiner\altaffilmark{1,2}}
\altaffiltext{1}{Department of Physics, Harvard University, 17 Oxford Street, 
Cambridge, MA 02138, USA; ameisner@fas.harvard.edu}
\altaffiltext{2}{Harvard-Smithsonian Center for Astrophysics, 60 Garden St, 
Cambridge, MA 02138, USA; dfinkbeiner@cfa.harvard.edu}

@@

We apply the \cite{FDS99} two-component thermal dust emission model to the
\PLANCK~HFI maps. This parametrization of the far-infrared dust spectrum as the
sum of two modified blackbodies serves as an important alternative to the 
commonly adopted single modified blackbody (MBB) dust emission model. Analyzing
the joint \PLANCK/DIRBE dust spectrum, we show that two-component models 
provide a better fit to the 100-3000 GHz emission than do single-MBB models, 
though by a lesser margin than found by \cite{FDS99} based on FIRAS and DIRBE. 
We also derive full-sky $6.1'$ resolution maps of dust optical depth and 
temperature by fitting the two-component model to the \PLANCK~217, 353, 545, 
and 857 GHz maps along with  DIRBE/\IRAS~100$\mu$m data. Because our 
two-component model matches the dust spectrum near its peak, accounts for the 
flattening of the dust spectrum at millimeter wavelengths, and specifies dust 
temperature at 6.1$'$ FWHM, our model provides reliable, high-resolution 
thermal dust emission foreground predictions from 100 GHz to 3000 GHz. We find 
that, in diffuse sky regions, our two-component 100-217 GHz thermal dust 
emission predictions are on average accurate to within $\le$2.2\%, which 
represents an improvement of 5-15\% relative to predictions based on the 
\cite{planckdust} single-MBB model. We calibrate our two-component optical 
depth to reddening at visible wavelengths, and compare the resulting reddening 
estimates with those based on stellar spectra. We find that imperfect zodiacal 
light subtraction is the dominant large-scale systematic problem in our 
two-component temperature map and reddening estimates.

@@

The primary FDS99 analysis considered the intrinsic grain properties of each
species, for example the emissivity power law indices, to be constant over the 
sky, and performed a correlation slope analysis to constrain these parameters 
with FIRAS and DIRBE observations. FDS99 also constructed a DIRBE 
240$\mu$m/100$\mu$m ratio to account for temperature variation at 
$\sim$1.3$^{\circ}$ resolution. In this work we seek to apply FDS99-like 
emission models to the \PLANCK~data set, which offers a dramatic enhancement in
angular resolution relative to the FIRAS spectra. The \PLANCK~data thereby 
allow us to derive an improved temperature correction at near-\IRAS~resolution
($\S$\ref{sec:mcmc}), re-evaluate the best-fit global dust properties 
($\S$\ref{sec:global}, $\S$\ref{sec:hier}), and fit additional parameters 
of the two-component model as a function of position on the sky 
($\S$\ref{sec:lores}). 
% in this last sentence reference the relevant sections for each sub-goal 
% mentioned

@@

\subsection{Single-MBB versus Two-component emission}
A major aim of this work has been to determine whether the  FDS99 
two-component dust emission model (or an FDS99-like model) remains favored over
single-MBB models when swapping the \PLANCK~HFI maps for FIRAS at frequencies 
below 1250 GHz. We compared dust SED models in two ways (1) by fitting a 
100-3000 GHz spectrum composed of per-band correlation slopes versus 
\PLANCK~857 GHz (2) by finding the best-fit dust temperature and optical depth
per line-of-sight, with each pixel's SED comprised of 100-3000 GHz 
\PLANCK+DIRBE data, and comparing the average goodness-of-fit under various
emission models.

>>>>

\title{Modeling thermal dust emission with two components: application to the 
{\it PLANCK} HFI maps}
\author{Aaron M. Meisner\altaffilmark{1,2}}
\author{Douglas P. Finkbeiner\altaffilmark{1,2}}
\altaffiltext{1}{Department of Physics, Harvard University, 17 Oxford Street, 
Cambridge, MA 02138, USA; ameisner@fas.harvard.edu}
\altaffiltext{2}{Harvard-Smithsonian Center for Astrophysics, 60 Garden St, 
Cambridge, MA 02138, USA; dfinkbeiner@cfa.harvard.edu}

@@

We apply the \cite{FDS99} two-component thermal dust emission model to the
\PLANCK~HFI maps. This parametrization of the far-infrared dust spectrum as the
sum of two modified blackbodies serves as an important alternative to the 
commonly adopted single modified blackbody (MBB) dust emission model. Analyzing
the joint \PLANCK/DIRBE dust spectrum, we show that two-component models 
provide a better fit to the 100-3000 GHz emission than do single-MBB models, 
though by a lesser margin than found by \cite{FDS99} based on FIRAS and DIRBE. 
We also derive full-sky $6.1'$ resolution maps of dust optical depth and 
temperature by fitting the two-component model to the \PLANCK~217, 353, 545, 
and 857 GHz maps along with  DIRBE/\IRAS~100$\mu$m data. Because our 
two-component model matches the dust spectrum near its peak, accounts for the 
flattening of the dust spectrum at millimeter wavelengths, and specifies dust 
temperature at 6.1$'$ FWHM, our model provides reliable, high-resolution 
thermal dust emission foreground predictions from 100 GHz to 3000 GHz. We find 
that, in diffuse sky regions, our two-component 100-217 GHz thermal dust 
emission predictions are on average accurate to within 2.2\%, which represents 
an improvement of 5-15\% relative to predictions based on the \cite{planckdust}
single-MBB model. We calibrate our two-component optical depth to reddening at 
visible wavelengths, and compare the resulting reddening estimates with those 
based on stellar spectra. We find imperfect zodiacal light subtraction in both 
DIRBE/\IRAS~and \PLANCK~to be the dominant large-scale systematic problem 
in our two-component temperature map and reddening estimates.

@@

The primary FDS99 analysis considered the intrinsic grain properties of each
species, for example the emissivity power law indices, to be constant over the 
sky, and performed a correlation slope analysis to constrain these parameters 
with FIRAS and DIRBE observations. FDS99 also constructed a DIRBE 
240$\mu$m/100$\mu$m ratio to account for temperature variation at 
$\sim$1.3$^{\circ}$ resolution. In this work we seek to apply the FDS99 
emission model to the \PLANCK~data set, which offers a dramatic enhancement in
angular resolution relative to the FIRAS spectra. The \PLANCK~data thereby 
allow us to derive an improved temperature correction at near-\IRAS~resolution
($\S$\ref{sec:mcmc}), re-evaluate the best-fit global dust properties 
($\S$\ref{sec:global}, $\S$\ref{sec:hier}), and fit additional 
two-component model parameters as a function of position on the sky 
($\S$\ref{sec:lores}).

@@

\subsection{Single-MBB versus Two-component emission}
A major aim of this work has been to determine whether the  FDS99 
two-component dust emission model remains favored over single-MBB models when 
swapping the \PLANCK~HFI maps for FIRAS at frequencies below 1250 GHz. We 
compared dust SED models in two ways (1) by fitting a 100-3000 GHz spectrum 
composed of per-band correlation slopes versus \PLANCK~857 GHz (2) by finding 
the best-fit dust temperature and optical depth per line-of-sight, with each 
pixel's SED comprised of 100-3000 GHz \PLANCK+DIRBE data, and comparing the 
average goodness-of-fit under various emission models.
